---
title: "Building a Personalized Recommendation System"
author: "Pascal Hermann Kouogang Tafo"
format: html
editor: visual
---

## INTRODUCTION

This assignment builds a personalized recommended system using survey data from "TV_Show_ratings.csv" that i manually created. The goal is to predict user preferences and recommend new TV shows, the users are likely to enjoy. To accomplish that goal, we will be implementing a **User-to-User Collaborative Filtering** algorithm using cosine similarity, generate top-N recommendations, and evaluate model performance.

------------------------------------------------------------------------

## APPROACH

The **User-to-User Collaborative Filtering** algorithm identifies users with similar tastes based on their past ratings and predicts ratings for items a user has not seen yet by **aggregating** the ratings of their "nearest neighbors". To build that personalized recommended system, we will be using the following approach:

-   Load the **"TV_Show_ratings.csv"** from my public GitHub and convert it into a wide-format user-item matrix.

-   Calculate the average rating for each user and subtract it from their ratings to account for individual rating biases.

-   Use **Cosine Similarity** on the centered ratings to determine how similar users are to one another.

-   For any unwatched show, calculate a weighted average of the ratings given by the most similar users.

------------------------------------------------------------------------

## Load the usefull library and the survey data

I will load the survey data that i created to build my personalized recommender system.

```{r}

# Loading Library

library(tidyverse)

# Loading the survey data

TV_Show_ratings <- read_csv ("https://raw.githubusercontent.com/Pascaltafo2025/Assignment-11-Personalized-Recommender-System/refs/heads/main/TV_Show_ratings.csv")

```

------------------------------------------------------------------------

## ALGORITHM CHOICE & IMPLEMENTATION

### Step 1: Create User-Item Matrix

```{r}

ratings_matrix <- TV_Show_ratings

ratings_matrix <- as.data.frame(ratings_matrix)
row.names(ratings_matrix) <- ratings_matrix$user_id
ratings_matrix$user_id <- NULL

ratings_matrix

```

### Step 2: Compute User Similarity (Cosine Similarity)

```{r}
cosine_sim <- function(x, y) {
  round(
    sum(x * y, na.rm = TRUE) / 
      (sqrt(sum(x^2, na.rm = TRUE)) * sqrt(sum(y^2, na.rm = TRUE))),
    2
  )
}

user_similarity <- matrix(
  NA,
  nrow = nrow(ratings_matrix),
  ncol = nrow(ratings_matrix)
)

for (i in 1:nrow(ratings_matrix)) {
  for (j in 1:nrow(ratings_matrix)) {
    user_similarity[i, j] <- cosine_sim(
      as.numeric(ratings_matrix[i, ]),
      as.numeric(ratings_matrix[j, ])
    )
  }
}

rownames(user_similarity) <- rownames(ratings_matrix)
colnames(user_similarity) <- rownames(ratings_matrix)

user_similarity

```

------------------------------------------------------------------------

### Step 3: Predict Missing Ratings and Generate Predictions

```{r}

# Predict Missing Ratings function

predict_rating <- function(user, item) {
  sims <- user_similarity[user, ]
  ratings <- ratings_matrix[, item]
  
  valid <- !is.na(ratings)
  
  sum(sims[valid] * ratings[valid]) / sum(abs(sims[valid]))
}

# Generate Predictions

predictions <- ratings_matrix

for(u in 1:nrow(ratings_matrix)){
  for(i in 1:ncol(ratings_matrix)){
    if(is.na(ratings_matrix[u, i])){
      predictions[u, i] <- round(predict_rating(u, i),2)
    }
  }
}

predictions

```

### Top Recommendation per User

Here, we will display recommended TV shows for a specific user and compute a practical example.

```{r}

get_top_N <- function(user_index, n=2) {
  user_ratings <- ratings_matrix[user_index, ]
  user_preds <- predictions[user_index, ]
  
  unseen <- is.na(user_ratings)
  
  top_items <- sort(user_preds[unseen], decreasing = TRUE)
  
  head(top_items, n)
}

# Example for user 1

get_top_N(1, 2)

```

------------------------------------------------------------------------

### Interpretation:

The recommender system outputs a ranked list of the top-N TV shows for each user that they have not yet rated, based on predicted ratings derived from similar users. Indeed, Alain who is the first user was recommended squid_game first and then the_witcher.

------------------------------------------------------------------------

## Model Evaluation

```{r}

# Convert from wide format to long format
ratings_long <- TV_Show_ratings %>%
  pivot_longer(
    cols = -user_id,
    names_to = "show",
    values_to = "rating"
  ) %>%
  filter(!is.na(rating))


set.seed(123)

ratings_split <- ratings_long %>%
  group_by(user_id) %>%
  mutate(test = sample(c(TRUE, FALSE), n(), replace = TRUE, prob = c(0.2, 0.8)))

train <- ratings_split %>% filter(test == FALSE)
test  <- ratings_split %>% filter(test == TRUE)

#RMSE Function

rmse <- function(actual, predicted) {
  sqrt(mean((actual - predicted)^2, na.rm = TRUE))
}
# Evaluate Model

test$predicted <- mapply(function(u, i) {
  predict_rating(u, i)
}, test$user_id, test$show)

# Calculation of Root Mean Squared Error (RMSE)

rmse_value <- rmse(test$rating, test$predicted)
rmse_value



```

## CONCLUSION

This assignment successfully developed a personalized recommender system using User-to-User Collaborative Filtering to predict user preferences and generate top-N TV show recommendations. By applying cosine similarity, predicting missing ratings, and evaluating performance using RMSE, the model demonstrated how recommendation systems can provide meaningful personalized suggestions from survey data.
